---
title: "lab10_warmup"
output: html_document
date: "2024-02-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Loading the Libraries:    

```{r}
library("tidyverse")
library("janitor")
```

### 1. In the data folder there is an epidemiology data set on an outbreak of malaria.    

```{r}
malaria <- read_csv("data/malaria.csv") %>% clean_names() #loading the data
```

### 2. `rdt` refers to rapid diagnostic test and they are identified here by age group.    

### 3. Make the data tidy and store them as a new object.      

```{r}
malaria
```

```{r}
 names(malaria)
```

```{r}
head(malaria) #evaluating tidyness of the data set
```

#### Tidying the data:    

```{r}
malaria_new <- malaria %>% 
  pivot_longer(cols=starts_with("malaria_rdt"),
               names_to = "age_class",
               values_to = "cases") %>% 
  select(newid, data_date, submitted_date, location_name, district, age_class, cases) 
malaria_new #note that this is still not perfect bc
```

```{r}
malaria %>% 
  pivot_longer(-c(location_name, data_date, submitted_date, province, district, malaria_tot, newid),
               names_to = "age_group",
               values_to = "malaria_rdt")
```


### 4. Which district had the highest total number of cases on July 30, 2020?      

```{r}
malaria_new %>% 
  filter(data_date == "2020-07-30") %>% 
  select(district, cases) %>% 
  group_by(district) %>% 
  summarize(total_cases = sum(cases, na.rm = T)) %>% 
  arrange(desc(total_cases)) #or arrange(-total_cases)
```

#### The district `Bolo` had the highest total number of cases on July 30, 2020.   
