---
title: "extra_practice_midterm2"
output: html_document
date: "2024-02-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(naniar)
library(gtools)
library(ggthemes)
```

```{r}
options(scipen=999)
```

# Lab 12 Practice   

```{r}
getwd()
```

```{r message=FALSE, warning=FALSE}
deserts <- read_csv("Desktop/BIS15W2024_cberevoescu/lab12/data/surveys_complete.csv")
```

```{r}
homerange <- read_csv("Desktop/BIS15W2024_cberevoescu/lab12/data/Tamburelloetal_HomeRangeDatabase.csv", na = c("", "NA", "\\"))
```
1. Use the `homerange` data to make a boxplot that shows the range of `log10.mass` by taxonomic class.
```{r}
homerange %>% 
        ggplot(aes(y = log10.mass, x = class)) +
        geom_boxplot() +
        labs(title = "Mass by Taxonomic Class",
             y = "Mass",
             x = "Class") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

2. Now, add a layer to this plot that fills color for each box by taxonomic class
```{r}
homerange %>% 
        ggplot(aes(y = log10.mass, x = class, fill = class)) +
        geom_boxplot() +
        labs(title = "Mass by Taxonomic Class",
             y = "Mass",
             x = "Class",
             fill = "Class") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

3. Use `group` to show the range of `log10.mass` by taxonomic class.
```{r}
homerange %>% 
        ggplot(aes(y = log10.mass, x = class, group = class)) +
        geom_boxplot() +
        labs(title = "Mass by Taxonomic Class",
             y = "Mass",
             x = "Class") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

4. Use `fill` to show the range of `log10.mass` by taxonomic class with color codes.
```{r}
homerange %>% 
        ggplot(aes(y = log10.mass, x = class, fill = class)) +
        geom_boxplot() +
        labs(title = "Mass by Taxonomic Class",
             y = "Mass",
             x = "Class",
             fill = "Class") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

1. Make a histogram of `log10.hra`. Make sure to add a title.
```{r}
homerange %>% 
        ggplot(aes(x = log10.hra)) +
        geom_histogram(color = "tomato3", fill = "tomato") +
        labs(title = "Distribution of Homerange",
             x = "Homerange",
             y = "Counts") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

2. Now plot the same variable using `geom_density()`.
```{r}
homerange %>% 
        ggplot(aes(x = log10.hra)) +
        geom_density(color = "tomato3", fill = "tomato", alpha = 0.5) +
        labs(title = "Distribution of Homerange",
             x = "Homerange",
             y = "Counts") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

3. Combine them both!
```{r}
homerange %>% 
        ggplot(aes(x = log10.hra)) +
        geom_histogram(aes(y = after_stat(density)), color = "tomato3", fill = "tomato") +
        geom_density(fill = "lavenderblush", color = "lavenderblush3", alpha = 0.5) +
        labs(title = "Distribution of Homerange",
             x = "Homerange",
             y = "Counts") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

1. Use case_when() to make a new column `range_category` that breaks down `log10.hra` into very small, small, medium, and large classes based on quartile.

```{r}
#install.packages("gtools")
library(gtools)
```

```{r}
quartiles <- quantcut(homerange$log10.hra)
table(quartiles)
```

```{r}
homerange %>% 
        mutate(range_category = case_when(log10.hra <= 3.65 ~ "very_small",
                                          log10.hra > 3.65 & log10.hra <= 4.59 ~ "small",
                                          log10.hra > 4.59 & log10.hra <= 6.02 ~ "medium",
                                          log10.hra > 6.02 ~ "large"))
```

2. Make a plot that shows how many and which taxonomic classes are represented in each `range_category`.   
```{r}
homerange %>% 
        mutate(range_category = case_when(log10.hra <= 3.65 ~ "very_small",
                                          log10.hra > 3.65 & log10.hra <= 4.59 ~ "small",
                                          log10.hra > 4.59 & log10.hra <= 6.02 ~ "medium",
                                          log10.hra > 6.02 ~ "large")) %>% 
        ggplot(aes(x = range_category, fill = class)) +
        geom_bar(position = "dodge") +
        theme_light() +
        scale_fill_brewer(palette = "Reds") +
        labs(title = "Taxonomic Classes Represented in Each Range Cateogory",
             x = "Range Category",
             y = "Count",
             fill = "Class") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

3. Isolate the small `range_category` and plot the range of `log10.mass` by taxonomic class.
```{r}
homerange %>% 
        mutate(range_category = case_when(log10.hra <= 3.65 ~ "very_small",
                                          log10.hra > 3.65 & log10.hra <= 4.59 ~ "small",
                                          log10.hra > 4.59 & log10.hra <= 6.02 ~ "medium",
                                          log10.hra > 6.02 ~ "large")) %>% 
        filter(range_category == "small") %>% 
        ggplot(aes(x = class, y = log10.mass, fill = class)) +
        geom_boxplot() +
        theme_bw() +
        scale_fill_brewer(palette = "RdPu") +
        labs(title = "Mass Range by Taxonomic Class",
             x = "Class",
             y = "Mass",
             fill = "Class") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

1. Build a scatterplot that compares `log10.mass` and `log10.preymass`. Color the points by `taxon`. Store this plot as object `q`.
```{r}
q <- homerange %>% 
        ggplot(aes(x = log10.mass, y = log10.preymass, color = taxon)) +
        geom_point(na.rm = T) +
        theme_gray() +
        labs(title = "Preymass vs Mass",
             x = "Mass",
             y = "Preymass",
             color = "Taxon") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5), legend.position = "left")
q
```

2. Use plot `q` but add the classic theme.
```{r}
q + theme_classic()
```

1. Use plot `q` and try the `theme_economist()` theme.
```{r}
q + theme_economist()
```

1. Keep using the `same `q` plot, but test out one of R Color Brewer palettes.
```{r}
q + scale_color_brewer(palette = "OrRd") 
```

1. Keep using the same `q` plot, but test out one of `paleteer` themes.
```{r}
q + scale_color_manual(values = paletteer_d("ggsci::nrc_npg"))
```

1. Use facet_wrap to compare the distribution of `log10.mass` by `thermoregulation`.
```{r}
homerange %>% 
        ggplot(aes(x = log10.mass)) +
        geom_histogram(color = "blue4", fill = "blue2") +
        facet_wrap(thermoregulation~.)
        
```

2. Use facet_grid to compare the distribution of `log10.mass` by `locomotion`.
```{r}
homerange %>% 
        ggplot(aes(x = log10.mass)) +
        geom_histogram(fill = "orange2", color = "orange3") +
        facet_grid(locomotion~.)
```

# Lab 11

For this practice, let's use the [palmerpenguins](https://allisonhorst.github.io/palmerpenguins/articles/intro.html) data.  

```{r}
library(palmerpenguins)
penguins <- penguins
```

1. Make a bar plot that shows counts of individuals by island. Fill by species, and try both a stacked bar plot and `position="dodge"`.
```{r}
penguins %>% 
        ggplot(aes(x = island, fill = species)) +
        geom_bar(position = "dodge") +
        scale_fill_brewer(palette = "Reds") +
        labs(title = "Individuals by Island",
             x = "Island",
             y = "Count") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

```{r}
penguins %>% 
        ggplot(aes(x = island, fill = species)) +
        geom_bar() +
        scale_fill_brewer(palette = "Reds") +
        labs(title = "Individuals by Island",
             x = "Island",
             y = "Count") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

2. Make another bar plot that shows the number of individuals by sex on each island?
```{r}
penguins %>% 
        filter(sex != "NA") %>% 
        ggplot(aes(x = island, fill = sex)) +
        geom_bar(position = "dodge") +
        scale_fill_brewer(palette = "Reds") +
        labs(title = "Individuals by Island",
             x = "Island",
             y = "Count") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

# Lab 10   

### 1. In the data folder there is an epidemiology data set on an outbreak of malaria.    

```{r}
malaria <- read_csv("Desktop/BIS15W2024_cberevoescu/lab10/data/malaria.csv") %>% clean_names() #loading the data
```

### 2. `rdt` refers to rapid diagnostic test and they are identified here by age group.    

### 3. Make the data tidy and store them as a new object.      

```{r}
malaria
```

```{r}
 names(malaria)
```

```{r}
head(malaria) #evaluating tidyness of the data set
```

#### Tidying the data:    

```{r}
malaria %>% 
        pivot_longer(cols = starts_with("malaria"),
                     names_to = "age_group",
                     values_to = "data")
```

### 4. Which district had the highest total number of cases on July 30, 2020?      

```{r}
malaria %>% 
        pivot_longer(cols = starts_with("malaria"),
                     names_to = "age_group",
                     values_to = "data") %>% 
        filter(data_date == "2020-7-30") %>% 
        group_by(district) %>% 
        summarize(total_data = sum(data, na.rm = T)) %>% 
        arrange(desc(total_data)) %>% 
        head(1)
```

#### The district `Bolo` had the highest total number of cases on July 30, 2020.   

```{r}
iris <- iris %>% clean_names()
```

1. Use the iris data to build a scatterplot that compares sepal length vs. sepal width. Use the cheat sheet for help to find the correct `geom_` for a scatterplot.   
```{r}
iris %>% 
        ggplot(aes(x = sepal_length, y = sepal_width)) +
        geom_point()
```

1. What is the structure of the `homerange` data? Does it have any NA's? Is it tidy? Do a quick exploratory analysis of your choice below.
```{r}
glimpse(homerange)
```

```{r}
homerange
```

#### It is tidy because each variable has its own column, each observation its own row, each value its own cell.    

```{r}
anyNA(iris)
```

```{r}
miss_var_summary(iris)
```

1. What is the relationship between log10.hra and log10.preymass? What do you notice about how ggplot treats NA's?    
```{r}
homerange %>% 
        ggplot(aes(x = log10.hra, y = log10.preymass)) +
        geom_point()
```

1. Filter the `homerange` data to include `mammals` only.
```{r}
homerange %>% 
        filter(class == "mammalia")
```

2. Are there more herbivores or carnivores in mammals? Make a bar plot that shows their relative numbers.

```{r}
homerange %>% 
        filter(class == "mammalia") %>% 
        ggplot(aes(x = trophic.guild)) +
        geom_bar()
```

3. Make a bar plot that shows the masses of the top 10 smallest mammals.
```{r}
homerange %>% 
        filter(class == "mammalia") %>% 
        arrange(mean.mass.g) %>% 
        head(10) %>% 
        ggplot(aes(x = species, y = mean.mass.g)) +
        geom_col()
```

1. What if we wanted a plot that showed the number of measured penguins for each species?
```{r}
penguins %>% 
        ggplot(aes(x = species)) +
        geom_bar()
```

2. How about average bill length by sex?
```{r}
penguins %>% 
        filter(sex != "NA") %>% 
        group_by(sex) %>% 
        summarize(average_bill_length = mean(bill_length_mm)) %>% 
        ggplot(aes(x = sex, y = average_bill_length)) +
        geom_col()
```


1. There are more herbivores than carnivores in the homerange data, but how do their masses compare? Make a summary and boxplot that compares their masses. Use `log10.mass`.   

```{r}
homerange %>% 
        group_by(trophic.guild) %>% 
        summarize(average_mass = mean(log10.mass, na.rm = T))
```

```{r}
homerange %>% 
        ggplot(aes(x = trophic.guild, y = log10.mass)) +
        geom_boxplot()
```

2. Have a closer look at carnivorous mammals. Summarize the range of log10.mass by family.
```{r}
homerange %>% 
        filter(trophic.guild == "carnivore", class == "mammalia") %>% 
        ggplot(aes(x = family, y = log10.mass)) +
        geom_boxplot() +
        coord_flip()
```

# Lab 9.   

1. Import the file `relig_income.csv` and store it as a new object `relig_income`.  
```{r}
relig_income <- read_csv("Desktop/BIS15W2024_cberevoescu/lab9/data/relig_income.csv")
```
2. Why are these data untidy?  

These data are untidy because the column names represent observations (income level).    

3. Use `pivot_longer()` to make the data tidy.  

```{r}
names(relig_income)
```

```{r}
relig_income %>% 
  pivot_longer(-religion,
               names_to = "income_level",
               values_to = "count")
```

1. Import `plant_data.csv` as a new object `plant_data`.  
```{r}
plant_data <- read_csv("Desktop/BIS15W2024_cberevoescu/lab9/data/plant_data.csv")
```

2. Why are these data not tidy?  

These data are not tidy because they have observations as column names.   

```{r}
names(plant_data)
```

3. Use `pivot_longer()` to make the data tidy. Focus the data only on genotype, water_sched_prog, and greenhouse.  
```{r}
plant_data %>% 
        pivot_longer(-c(genotype, water_sched_prog, greenhouse),
                     names_to = "day",
                     values_to = "data")
```

1. Are these data tidy? (NO) Please use `pivot_longer()` to tidy the data.

```{r}
expenditures <- USPersonalExpenditure %>% 
  as_tibble() %>% #this transforms the matrix into a data frame
  mutate(expenditure = rownames(USPersonalExpenditure))
expenditures
```

```{r}
names(expenditures)
```

```{r}
expenditures %>% 
        pivot_longer(-expenditure,
                     names_to = "year",
                     values_to = "data")
```

2. Restrict the data to medical and health expenditures only. Sort in ascending order.
```{r}
expenditures %>% 
        pivot_longer(-expenditure,
                     names_to = "year",
                     values_to = "data") %>% 
        filter(expenditure == "Medical and Health") %>% 
        arrange(data)
```

3. Restrict the data to private education only. Sort in ascending order.   
```{r}
expenditures %>% 
        pivot_longer(-expenditure,
                     names_to = "year",
                     values_to = "data") %>% 
        filter(expenditure == "Private Education") %>% 
        arrange(data)
```

```{r}
heartrate2 <- read_csv("Desktop/BIS15W2024_cberevoescu/lab9/data/heartrate2.csv")
heartrate2 %>% 
  separate(patient, into= c("patient", "sex"), sep = "_")
```
1. Re-examine `heartrate2`. Use `separate()` for the sexes, `pivot_longer()` to tidy, and `arrange()` to organize by patient and drug. Store this as a new object `heartrate3`.  
```{r}
heartrate3 <- heartrate2 %>% 
        separate(patient, into = c("patient", "sex"), sep = "_") %>% 
        pivot_longer(-c(patient, sex),
                     names_to = "treatment",
                     values_to = "heartrate")
heartrate3
```


2. `unite()` is the opposite of separate(). Its syntax is straightforward. You only need to give a new column name and then list the columns to combine with a separation character.  Give it a try below by recombining patient and sex from `heartrate3`.  
```{r}
heartrate3 %>% 
        unite(patient_sex, "patient", "sex", sep = "_")
```

1. Load the `gene_exp.csv` data as a new object `gene_exp`. Are these data tidy? Use `pivot_wider()` to tidy the data.    

No, there are more than one variable represented in each column.    

```{r}
gene_exp <- read_csv("Desktop/BIS15W2024_cberevoescu/lab9/data/gene_exp.csv")
gene_exp
```
```{r}
names(gene_exp)
```

```{r}
gene_exp %>% 
  pivot_wider(names_from = type,
              values_from = L4_values)
```

## Practice
For the last practice example, I will use data from the awesome [R Ladies Sydney](https://rladiessydney.org/courses/ryouwithme/02-cleanitup-5/) blog. This data set is compiled by the NSW Office of Environment and Heritage contains the enterococci counts in water samples obtained from Sydney beaches as part of the Beachwatch Water Quality Program! The data set we’ll be working with is current as of October 13th 2018.  

1. Load the beachbugs data and have a look.
```{r}
beachbugs <- read_csv("Desktop/BIS15W2024_cberevoescu/lab9/data/beachbugs_long.csv")
beachbugs
```

```{r}
beachbugs
```

2. Use `pivot_wider` to transform the data into wide format.   
```{r}
beachbugs_wide <- beachbugs %>% 
        pivot_wider(names_from = site,
                    values_from = buglevels)
beachbugs_wide
```

3. Now, use `pivot_longer` to transform them back to long!
```{r}
beachbugs
```

```{r}
beachbugs_wide %>% 
        pivot_longer(-year,
                     names_to = "site",
                     values_to = "bug_level")
```

# Lab 8.   

1. Produce a summary of the mean for bill_length_mm, bill_depth_mm, flipper_length_mm, and body_mass_g within Adelie penguins only. Be sure to provide the number of samples.
```{r}
penguins
```

```{r}
penguins %>% 
        filter(species == "Adelie") %>%
        group_by(species) %>% 
        summarize(across(contains("mm"), mean, na.rm = T),
                  body_mass_g = mean(body_mass_g, na.rm = T))
```

2. How does the mean of `bill_length_mm` compare between penguin species?
```{r}
penguins %>% 
       group_by(species) %>% 
        summarise(average_bill_length_mm = mean(bill_length_mm, na.rm = T))
```

3. For some penguins, their sex is listed as NA. Where do these penguins occur?
```{r}
penguins %>% 
        count(sex, island)
```

1. Produce separate summaries of the mean and standard deviation for bill_length_mm, bill_depth_mm, and flipper_length_mm for each penguin species. Be sure to provide the number of samples.  
```{r}
penguins %>% 
        group_by(species) %>% 
        summarize(across(contains("mm"), mean, na.rm = T),
                  n_samples = n())
```

1. Can we use a single approach to deal with NA's in this data set? Given what you learned in the previous lab, how would you manage the NA values?
```{r}
life_history %>% 
        map_df(~ sum(is.na(.)))
```

```{r}
miss_var_summary(life_history)
```

1. Import the data and do a little exploration. Be sure to clean the names if necessary.
```{r}
cites_wildlife <- read_csv("Desktop/BIS15W2024_cberevoescu/lab8/data/cites.csv") %>% clean_names()
```

```{r}
summary(cites_wildlife)
```

2. Use `naniar` to summarize the NA's in each variable.
```{r}
miss_var_summary(cites_wildlife)
```

3. Try using `group_by()` with `naniar`. Look specifically at class and `exporter_reported_quantity`. For which taxonomic classes do we have a high proportion of missing export data?
```{r}
cites_wildlife %>% 
        group_by(class) %>% 
        select(class, exporter_reported_quantity) %>% 
        miss_var_summary() %>% 
        arrange(desc(pct_miss))
```

# Midterm 2 Practice.   

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your code should be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean! Use the tidyverse and pipes unless otherwise indicated. To receive full credit, all plots must have clearly labeled axes, a title, and consistent aesthetics. This exam is worth a total of 35 points. 

Please load the following libraries.
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
library("naniar")
```

## Data
These data are from a study on surgical residents. The study was originally published by Sessier et al. “Operation Timing and 30-Day Mortality After Elective General Surgery”. Anesth Analg 2011; 113: 1423-8. The data were cleaned for instructional use by Amy S. Nowacki, “Surgery Timing Dataset”, TSHS Resources Portal (2016). Available at https://www.causeweb.org/tshs/surgery-timing/.

Descriptions of the variables and the study are included as pdf's in the data folder.  

Please run the following chunk to import the data.
```{r message=FALSE, warning=FALSE}
surgery <- read_csv("Desktop/BIS15W2024_cberevoescu/practice midterm2/data/surgery.csv")
```

1. Use the summary function(s) of your choice to explore the data and get an idea of its structure. Please also check for NA's.
```{r}
glimpse(surgery)
```

```{r}
miss_var_summary(surgery)
```

2. Let's explore the participants in the study. Show a count of participants by race AND make a plot that visually represents your output.
```{r}
surgery %>% 
        count(race)
```

```{r}
surgery %>% 
        ggplot(aes(x = race, fill = race)) + 
        geom_bar() +
        theme_economist() +
        scale_fill_brewer(palette = "Blues") +
        labs(title = "Participant Race",
             x = "Race",
             y = "Count") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5), legend.position = "right")
```


3. What is the mean age of participants by gender? (hint: please provide a number for each) Since only three participants do not have gender indicated, remove these participants from the data.
```{r}
surgery %>% 
        filter(gender != "NA") %>% 
        group_by(gender) %>% 
        summarize(mean_age = mean(age, na.rm = T))
```

The mean age of female participants is 56.70730 years, and the mean age of male participants is 58.77323 years.   

4. Make a plot that shows the range of age associated with gender.
```{r}
surgery %>% 
        filter(gender != "NA") %>% 
        ggplot(aes(x = gender, y = age)) +
        geom_boxplot(na.rm = T) +
        theme_economist() +
        scale_fill_brewer(palette = "Blues") +
        labs(title = "Age by Participant Gender",
             x = "Gender",
             y = "Age") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

5. How healthy are the participants? The variable `asa_status` is an evaluation of patient physical status prior to surgery. Lower numbers indicate fewer comorbidities (presence of two or more diseases or medical conditions in a patient). Make a plot that compares the number of `asa_status` I-II, III, and IV-V.
```{r}
surgery %>% 
        filter(asa_status != "NA") %>% 
        ggplot(aes(x = asa_status, fill = asa_status)) +
        geom_bar() +
        theme_economist() +
        scale_fill_brewer(palette = "Blues") +
        labs(title = "Health of Participants",
             x = "ASA Status",
             y = "Count") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

6. Create a plot that displays the distribution of body mass index for each `asa_status` as a probability distribution- not a histogram. (hint: use faceting!)
```{r}
surgery %>% 
        filter(asa_status != "NA") %>% 
        ggplot(aes(x = bmi)) +
        geom_density(na.rm = T, color = "tomato", fill = "tomato3") +
        facet_wrap(~asa_status)
        theme_economist() +
        labs(title = "Distribution of BMI for ASA Status",
             x = "BMI",
             y = "Count") +
        theme(plot.title = element_text(size = rel(1.3), hjust = 0.5))
```

The variable `ccsmort30rate` is a measure of the overall 30-day mortality rate associated with each type of operation. The variable `ccscomplicationrate` is a measure of the 30-day in-hospital complication rate. The variable `ahrq_ccs` lists each type of operation.  

7. What are the 5 procedures associated with highest risk of 30-day mortality AND how do they compare with the 5 procedures with highest risk of complication? (hint: no need for a plot here)
```{r}
surgery %>% 
        group_by(ahrq_ccs) %>% 
        summarise(average_mortality = mean(ccsmort30rate, na.rm = T)) %>% 
        arrange(desc(average_mortality)) %>% 
        head(5)
```

```{r}
surgery %>% 
        group_by(ahrq_ccs) %>% 
        summarise(average_complication = mean(ccscomplicationrate, na.rm = T)) %>% 
        arrange(desc(average_complication)) %>% 
        head(5)
```

There is a lot of overlap between the two lists.    

8. Make a plot that compares the `ccsmort30rate` for all listed `ahrq_ccs` procedures.
```{r}
surgery %>% 
        ggplot(aes(x = ahrq_ccs, y = ccsmort30rate, fill = ahrq_ccs)) +
        geom_col() +
        coord_flip() +
        theme_economist() +
        labs(title = "Mortality Rate Compared for All Procedures",
             x = "Procedure",
             y = "Mortality Rate") +
        theme(plot.title = element_text(size = rel(2.0), hjust = 0.5), legend.position = "bottom")
```

9. When is the best month to have surgery? Make a chart that shows the 30-day mortality and complications for the patients by month. `mort30` is the variable that shows whether or not a patient survived 30 days post-operation.
```{r}
surgery %>% 
        mutate(mort_30 = ifelse(mort30 == "Yes", 1, 0),
               comp = ifelse(complication == "Yes", 1, 0)) %>% 
        group_by(month) %>% 
        summarize(month_mort_30 = sum(mort_30, na.rm = T),
                  month_comp = sum(comp, na.rm = T)) %>% 
        arrange(month_mort_30)
```

It appears that December is the best month to have surgery, because it is the month with the lowest number of mortalities and complications.   

10. Make a plot that visualizes the chart from question #9. Make sure that the months are on the x-axis. Do a search online and figure out how to order the months Jan-Dec.
```{r}
surgery %>% 
        mutate(mort_30 = ifelse(mort30 == "Yes", 1, 0),
               comp = ifelse(complication == "Yes", 1, 0)) %>% 
        group_by(month) %>% 
        summarize(month_mort_30 = sum(mort_30, na.rm = T),
                  month_comp = sum(comp, na.rm = T)) %>% 
        ggplot(aes(x = month, y = month_mort_30, fill = month_comp)) +
        geom_col(alpha = 0.75) +
        scale_x_discrete(limits = month.abb)
```

Please be 100% sure your exam is saved, knitted, and pushed to your github repository. No need to submit a link on canvas, we will find your exam in your repository.
