---
title: "BIS15L_AI_hmwk"
author: "Catrinel Berevoescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions   

This homework is a bit less traditional this week. There are some errors in the code below. In addition add notes and titles to the code chunks that explains what the code chunck does. Use AI tools, notes from class, or any other resources to help fix the errors. Feel free to edit any plots to your preference. Don't forget to answer the two questions at the bottom. Upon completion, you should not have to '#' out any of the code chunks.

Push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.   

Again make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

```{r}
library(tidyverse) #loading the libraries
library(janitor)
library(dplyr)
```

### 1. Uploading the data set `blood_count_dataset.csv`.   

```{r}
blood_count <- read_csv("data/blood_count_dataset.csv") %>% clean_names() #uploading the data set, and cleaning the variable names
```

### 2. Getting a Summary of the data set `blood_count`.   

```{r}
glimpse(blood_count) #using the glimpse() function to get a summary of the `blood-count` data set
```

### 3. Finding the minimum, maximum, and mean of the red blood cell counts by gender.  

```{r}
blood_count %>% 
  group_by(gender) %>% #grouping by the variable `gender`
  summarize(min = min(red_blood_cells), #generating a statistical summary of the `red_blood_cells` variable by gender
            mean = mean(red_blood_cells),
            max = max(red_blood_cells))
```

### 4. Finding the mean value of columns 3-9 for males.     

```{r}
options(scipen = 999, digits = 2) #canceling the use of scientific notation
```

```{r}
blood_count %>% 
  filter(gender == "Male") %>% #filtering the data set by only males
  select(3:9) %>% #selecting columns 3 through 9
  colMeans(na.rm = TRUE) #finding column means, removing NA's
```

### 5. Finding the mean value of columns 3-9 for females.     

```{r}
blood_count %>% 
  filter(gender == "Female")  %>% #filtering the data set by only females
  select(3:9) %>% #selecting columns 3 through 9
  colMeans(na.rm = T) #finding column means, removing NA's
```

### 6. Creating boxplots of the blood count tests values by sex.   

```{r}
blood_count %>%
  pivot_longer(-c(age, gender), #pivoting the data set to a long format, with blood count test type in one column and values in another
               names_to = "test", 
               values_to = "values") %>%
  ggplot(aes(x = gender, y = values, fill = gender)) + #creating the plot
  geom_boxplot() + #creating a box plot
  facet_wrap(~test) + #using facet_wrap() with the `test` variable for ease of viewing
  labs(title = "Boxplot of Blood Count Tests",
       x = "Test",
       y = "Values",
       fill = "Gender") + #adding labels
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5)) #changing the title size and location
```

### 7. Creating a new object with a new age group column.   

```{r}
grouped_bc <- blood_count %>% #creating a new object with a new column for age group
  mutate(age_group = case_when(
    age >= 20 & age <= 29 ~ "20s",
    age >= 30 & age <= 39 ~ "30s",
    age >= 40 & age <= 49 ~ "40s",
    age >= 50 & age <= 59 ~ "50s",
    age >= 60 & age <= 70 ~ "60s")) #there were only 4 people in 70s, so I combined it with 60s
head(grouped_bc) #checking the first few rows of the new object
```

### 8. Creating a Boxplot of Blood Count Tests by Age Group.  

```{r}
grouped_bc %>%
  pivot_longer(3:9, 
               names_to = "test", 
               values_to = "values") %>% #pivoting the dataset to a long format, with blood count test type in one column and values in another
  ggplot(aes(x = age_group, y = values, fill = age_group)) + #creating the plot
  geom_boxplot() + #creating a box plot
  facet_wrap(~test, scales = "free") +  #using facet_wrap() with the `test` variable
  labs(title = "Boxplot of Blood Count Tests by Age Group",
       x = "Age Group",
       y = "Values",
       fill = "Age Group") + #adding labels
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5)) #adjusting the size and location of the title
```

### 9. What are two drawbacks and two benefits of AI?

#### Two drawbacks to using AI are that (1) code created using AI can be messy and not totally accurate, while appearing to be so (and can therefore be misleading), and (2) the resulting broken code can be difficult to troubleshoot.   
#### Two benefits of AI are that (1) it can be used to automate repetitive and simple tasks, and (2) it can be used to deal with large numbers of big data sets quickly and efficiently.   

### 10. Do you think you will use AI for coding in the future? Why or why not?   

#### I do not think I will be using AI for coding in the future, because I find troubleshooting to be the more challenging part of coding, and I think that AI can actually distract from that process. I may use it for simple/shorter code, but for creating large chunks of code, I am finding that AI can create issues pretty quickly.   

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences. 