---
title: "BIS15L_AI_hmwk"
author: "Catrinel Berevoescu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions   

This homework is a bit less traditional this week. There are some errors in the code below. In addition add notes and titles to the code chunks that explains what the code chunck does. Use AI tools, notes from class, or any other resources to help fix the errors. Feel free to edit any plots to your preference. Don't forget to answer the two questions at the bottom. Upon completion, you should not have to '#' out any of the code chunks.

Push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.   

Again make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

```{r}
library(tidyverse)
library(janitor)
library(dplyr)
```

### 1. Uploading the data set `blood_count_dataset.csv`.   

```{r}
blood_count <- read_csv("data/blood_count_dataset.csv") %>% clean_names() #ensuring the dataset has the correct name
```

### 2. Getting a Summary of the data set.   

```{r}
glimpse(blood_count)
```

### 3. Finding the minimum, maximum, and mean of the red blood cell counts by gender.  

```{r}
blood_count %>% 
  group_by(gender) %>% 
  summarize(min = min(red_blood_cells),
            mean = mean(red_blood_cells),
            max = max(red_blood_cells))
```

### 4. Finding the mean value of columns 3-9 for males.     

```{r}
options(scipen = 999, digits = 2)
```

```{r}
blood_count %>% 
  filter(gender == "Male") %>%
  select(3:9) %>%
  colMeans(na.rm = TRUE)
```

### 5. Finding the mean value of columns 3-9 for females.     

```{r}
blood_count %>% 
  filter(gender == "Female")  %>% 
  select(3:9) %>%
  colMeans(na.rm = T)
```

### 6. Creating boxplots of the blood count tests values by sex.   

```{r}
blood_count %>%
  pivot_longer(-c(age, gender), 
               names_to = "test", 
               values_to = "values") %>%
  ggplot(aes(x = gender, y = values, fill = gender)) +
  geom_boxplot() +
  facet_wrap(~test) +
  labs(title = "Boxplot of Blood Count Tests",
       x = "Test",
       y = "Values",
       fill = "Gender") +
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```

### 7. Creating a new object with a new age group column.   

```{r}
grouped_bc <- blood_count %>%
  mutate(age_group = case_when(
    age >= 20 & age <= 29 ~ "20s",
    age >= 30 & age <= 39 ~ "30s",
    age >= 40 & age <= 49 ~ "40s",
    age >= 50 & age <= 59 ~ "50s",
    age >= 60 & age <= 70 ~ "60s")) # there were only 4 people in 70s, so I combined it with 60s
head(grouped_bc)

```

### 8. Creating a Boxplot of Blood Count Tests by Age Group.  

```{r}
grouped_bc %>%
  pivot_longer(3:9, names_to = "test", values_to = "values") %>%
  ggplot(aes(x = age_group, y = values, fill = age_group)) +
  geom_boxplot() +
  facet_wrap(~test, scales = "free") +  # Facet by the "test" variable
  labs(title = "Boxplot of Blood Count Tests by Age Group",
       x = "Age Group",
       y = "Values",
       fill = "Age Group") +
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```

### 9. What are two drawbacks and two benefits of AI?

#### Two drawbacks to using AI are that code created using AI can be messy and not totally accurate, while appearing to be so (can be misleading), and the resulting broken code can be difficult to troubleshoot. Two benefits of AI are that it can be used to automate repetitive and simple tasks, and it can be used to deal with large numbers of big data sets quickly and efficiently.   

### 10. Do you think you will use AI for coding in the future? Why or why not?   

#### I do not think I will be using AI for coding in the future, because I find troubleshooting to be the more challenging part of coding, and I think that AI can actually distract from that process. I may use it for simple/shorter code, but for creating large chunks of code, I am finding that AI can create issues pretty quickly.   

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences. 