---
title: "Homework 7"
author: "Catrinel Berevoescu"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
library(naniar)
``` 

## Data   

### 1. For this homework, we will use two different data sets. Please load `amniota` and `amphibio`.  

`amniota` data:  
Myhrvold N, Baldridge E, Chan B, Sivam D, Freeman DL, Ernest SKM (2015). “An amniote life-history
database to perform comparative analyses with birds, mammals, and reptiles.” _Ecology_, *96*, 3109.
doi: 10.1890/15-0846.1 (URL: https://doi.org/10.1890/15-0846.1).    

```{r}
amniota <- read_csv("data/amniota.csv") %>% clean_names()
```

`amphibio` data:  
Oliveira BF, São-Pedro VA, Santos-Barrera G, Penone C, Costa GC (2017). “AmphiBIO, a global database
for amphibian ecological traits.” _Scientific Data_, *4*, 170123. doi: 10.1038/sdata.2017.123 (URL:
https://doi.org/10.1038/sdata.2017.123).   

```{r}
amphibio <- read_csv("/Users/catrinelberevoescu/Desktop/BIS15W2024_cberevoescu/lab8/data/amphibio.csv") %>% clean_names()
```

## Questions    

### 2. Do some exploratory analysis of the `amniota` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.    

```{r}
glimpse(amniota)
```

#### NA's are represented in the data set `amniota` with the values -999, -999.0, -999.00, -999.000, -999.0000, -999.00000.    

How Many NA's are Represented in `amniota`:    

```{r}
miss_var_summary(amniota)
```

### 3. Do some exploratory analysis of the `amphibio` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.  

```{r}
glimpse(amphibio) #this shows that the variable fruits is logical
```

```{r}
summary(amphibio) #this indicates that 0s COULD have been used as "NA's" for the variables dir, lar, viv, considering that the data in these columns is binary as in other columns in this data set
```

#### There are no strange/outlandish values that could clearly have been used as placeholders for NA's in this dataset. It seems that NA's are represented by NA in the `amphibio` data set.    

How Many NA's are Represented in `amphibio`:    

```{r}
miss_var_summary(amphibio)
```

### 4. How many total NA's are in each data set? Do these values make sense? Are NA's represented by values?    

`amniota` Data Set:   

```{r}
amniota %>% 
  summarise(total_na = sum(is.na(.))) #total NAs
```

#### There are NO NAs in the `amniota` data set. The fact that there are NO NAs in the `amniota` data set does NOT make sense. When using the summary() function (below), it DOES seem that there are NA's represented by values (-999, -999.0, -999.000):    

```{r}
summary(amniota)
```

`amphibio` Data Set:   

```{r}
amphibio %>% 
  summarise(total_na = sum(is.na(.)))
```

#### There are ARE NAs in the `amphibio` data set, 170566 in total. The fact that there are NAs in the `amphibio` data set does make sense. When using the summary() function (below), it does NOT seem that there are NA's represented by placeholder values:    

```{r}
summary(amphibio)
```

### 5. Make any necessary replacements in the data such that all NA's appear as "NA".  

`amniota` Data Set:    

```{r}
amniota <- amniota %>% 
  replace_with_na_all(condition = ~.x == -999) 
amniota
```

#### I determined that there are no clear placeholder values for NA's in the `amphibio` data set - it seems that all the NA's are already represented by NA's.    

### 6. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amniota` data.  

```{r}
miss_var_summary(amniota) #summary with percentages of missing data for each variable in amniota data
```

### 7. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amphibio` data.    

```{r}
miss_var_summary(amphibio) #summary with percentages of missing data for each variable in amphibio data
```

### 8. For the `amniota` data, calculate the number of NAs in the `egg_mass_g` column sorted by taxonomic class; i.e. how many NA's are present in the `egg_mass_g` column in birds, mammals, and reptiles? Does this results make sense biologically? How do these results affect your interpretation of NA's?    

```{r}
amniota %>%
  group_by(class) %>%
  select(class, egg_mass_g) %>% 
  miss_var_summary() %>% 
  arrange(desc(pct_miss))
```

#### These results make SOME sense biologically. Mammals are shown to have 100% missing `egg_mass_g values`, which DOES make sense because the vast majority of Mammals do not lay eggs. The large number of NA's for the Aves (birds) and Reptilia (Reptiles) makes LESS sense biologically because most birds and many reptiles lay eggs.    
#### These results indicate that some of the NA's exist because of biological reasons and not because of issues with measurement taking - thus changing my interpretation of some of the NA's.    

### 9. The `amphibio` data have variables that classify species as fossorial (burrowing), terrestrial, aquatic, or arboreal. Calculate the number of NA's in each of these variables. Do you think that the authors intend us to think that there are NA's in these columns or could they represent something else? Explain.     

Calculating the number of NA's in the fossorial (burrowing), terrestrial, aquatic, or arboreal variables:   

```{r}
amphibio %>%
  select(fos, ter, aqu, arb) %>% 
  miss_var_summary() %>% 
  arrange(desc(pct_miss))
```

#### There are 6053 NA's in the fossorial variable, 4347 NA's in the arboreal variable, 2810 NA's in the aquatic variable, and 1104 NA's in the terrestrial variable.    
#### Since these variables are used to categorize the species, I believe that the authors intended to indicate which categories that a certain species is part of (1) OR not (NA). The NA's could represent that the species is NOT part of that variable's category.    

### 10. Now that we know how NA's are represented in the `amniota` data, how would you load the data such that the values which represent NA's are automatically converted?     

Loading the data such that the values that represent NA's are automatically converted (nuclear option):     

```{r}
amniota_new <- read_csv(file = "data/amniota.csv", na=c("-999")) %>% 
  clean_names()
```

```{r}
amniota_new
```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  