---
title: "BIS15L_AI_class"
author: "Catrinel Berevoescu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?  

Github Copilot Response: AI is the simulation of human intelligence processes by machines, especially computer systems. These processes include learning, reasoning, and self-correction.  

2. What is ChatGPT?   

Github Copilot Response: ChatGPT is a language model developed by OpenAI that can generate human-like text based on the input it receives. It is a powerful tool for generating text and can be used for a variety of tasks, such as writing code, answering questions, and generating creative writing.    

3. What is Github Copilot?   

Github Copilot Response: Github Copilot is an AI-powered code completion tool that is designed to help developers write code more quickly and efficiently. It uses machine learning to analyze the code you are writing and suggest completions based on the context. It can be a helpful tool for writing code, but it is important to use it as a supplement to your own knowledge and not as a replacement for understanding the code you are writing.  

4. [Try it yourself: Insert your own question here].  Why should I use AI tools in R?   

Github Copilot Response: I should use AI tools in R because they can help me write code more quickly and efficiently. They can also help me troubleshoot code and find solutions to problems. However, it is important to use AI tools as a supplement to my own knowledge and not as a replacement for understanding the code I am writing.  

Will get different responses based off of factors such as previous experience and the computer you use (it is responsive). 

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix
netflix <- read_csv("Data/netflix_titles.csv") 
```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*

Yes it did, it was a bit creepy honestly.   

## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
#use names() to get the column names of the data
names(netflix)
```

```{r}
summary(netflix) #get the summary of the data
```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*

Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
netflix_movies <- netflix %>% 
  filter(type == "Movie") #filter the data to only include movies
```

Which director has the most movies on netflix?
```{r}
netflix_movies %>% 
  count(director, sort = TRUE) #count the number of movies by director
```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*

[insert your prompt here] Which director has the most movies, excluding NA's, on netflix?
```{r}
netflix_movies %>% 
  count(director, sort = TRUE) %>% 
  filter(!is.na(director)) %>% #count the number of movies by director
  arrange(desc(n)) #arrange the data by the number of movies
```

The director that has the most movies on netflix is Rajiv Chilaka.   

Which director had the most movies by release year and remove the NA's?
```{r}
netflix_movies %>% 
  count(director, release_year, sort = TRUE) %>% 
  filter(!is.na(director)) %>% #count the number of movies by director and release year
  arrange(desc(n)) #arrange the data by the number of movies
```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


Which individual directors had the most movies by release year, remove the NA's, and separate by director name, and rearrange data?
```{r}
netflix_movies %>% 
  count(director, release_year, sort = TRUE) %>% 
  filter(!is.na(director)) %>% #count the number of movies by director and release year
  separate(director, sep = ", ", into = c("director1", "director2")) %>% #separate the director names
  arrange(desc(n)) #arrange the data by the number of movies
```

```{r}
netflix_movies %>% 
  count(director, release_year, sort = TRUE) %>% 
  filter(!is.na(director)) %>% #count the number of movies by director and release year
  separate(director, sep = ", ", into = c("director1", "director2")) %>% #separate the director names
  arrange(desc(n)) #arrange the data by the number of movies
```

Separate and add movies by each director for each year into their own row
```{r}
netflix_movies %>% 
  count(director, release_year, sort = TRUE) %>% 
  filter(!is.na(director)) %>% #count the number of movies by director and release year
  separate(director, sep = ", ", into = c("director1", "director2")) %>% #separate the director names
  pivot_longer(cols = starts_with("director"), values_to = "director") %>% #separate each director into their own row
  arrange(desc(n)) #arrange the data by the number of movies
```

```{r}
netflix_movies %>% 
  count(director, release_year, sort = TRUE) %>% 
  filter(!is.na(director)) %>% #count the number of movies by director and release year
  separate(director, sep = ", ", into = c("director1", "director2")) %>% #separate the director names
  arrange(desc(n)) %>% #arrange the data by the number of movies
  pivot_wider(names_from = release_year, values_from = n, values_fill = 0) #separate and add movies by each director for each year into their own row
```

Make a bar plot of the top directors with the most movies on netflix. 
```{r}
netflix_movies %>% 
  count(director, sort = TRUE) %>% 
  filter(!is.na(director)) %>% #count the number of movies by director
  arrange(desc(n)) %>% #arrange the data by the number of movies
  slice(1:10) %>% #select the top 10 directors
  ggplot(aes(x = reorder(director, n), y = n)) + #create a bar plot
  geom_bar(stat = "identity") + #use the count of movies as the y variable
  labs(title = "Top 10 Directors with the Most Movies on Netflix", x = "Director", y = "Number of Movies") + #add a title and axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) #rotate the x axis labels
```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

Make a neat, clear, colorful bar plot of the top directors with the most movies on netflix
```{r}
netflix_movies %>% 
  count(director, sort = TRUE) %>% 
  filter(!is.na(director)) %>% #count the number of movies by director
  arrange(desc(n)) %>% #arrange the data by the number of movies
  slice(1:10) %>% #select the top 10 directors
  ggplot(aes(x = reorder(director, n), y = n, fill = director)) + #create a bar plot
  geom_bar(stat = "identity") + #use the count of movies as the y variable
  labs(title = "Top 10 Directors with the Most Movies on Netflix", x = "Director", y = "Number of Movies") + #add a title and axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) #rotate the x axis labels
```

*How about for just 2020*

Make a neat, clear, colorful bar plot of the top directors with the most movies on netflix in 2020
```{r}
netflix_movies %>% 
  filter(release_year == 2020) %>% #filter the data to only include movies from 2020
  count(director, sort = TRUE) %>% 
  filter(!is.na(director)) %>% #count the number of movies by director
  arrange(desc(n)) %>% #arrange the data by the number of movies
  slice(1:10) %>% #select the top 10 directors
  ggplot(aes(x = reorder(director, n), y = n, fill = director)) + #create a bar plot
  geom_bar(stat = "identity") + #use the count of movies as the y variable
  labs(title = "Top 10 Directors with the Most Movies on Netflix in 2020", x = "Director", y = "Number of Movies") + #add a title and axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) #rotate the x axis labels
```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*

## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt

Create an interactive shiny app that displays a neat, clear, colorful bar plot of the top 10 directors with the most movies on netflix
```{r}
ui <- fluidPage(
  titlePanel("Top 10 Directors with the Most Movies on Netflix"),
  plotOutput("barplot")
)
server <- function(input, output) {
  output$barplot <- renderPlot({
    netflix_movies %>% 
      count(director, sort = TRUE) %>% 
      filter(!is.na(director)) %>% #count the number of movies by director
      arrange(desc(n)) %>% #arrange the data by the number of movies
      slice(1:10) %>% #select the top 10 directors
      ggplot(aes(x = reorder(director, n), y = n, fill = director)) + #create a bar plot
      geom_bar(stat = "identity") + #use the count of movies as the y variable
      labs(title = "Top 10 Directors with the Most Movies on Netflix", x = "Director", y = "Number of Movies") + #add a title and axis labels
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) #rotate the x axis labels
  })
}
shinyApp(ui = ui, server = server)
```

Make it more interactive.  
```{r}
ui <- fluidPage(
  titlePanel("Top 10 Directors with the Most Movies on Netflix"),
  selectInput("x", "Select Year", choices = unique(netflix_movies$release_year), hr()),
  plotOutput("barplot")
)
server <- function(input, output) {
  output$barplot <- renderPlot({
    netflix_movies %>% 
      filter(release_year == input$x) %>% #filter the data to only include movies from the selected year
      count(director, sort = TRUE) %>% 
      filter(!is.na(director)) %>% #count the number of movies by director
      arrange(desc(n)) %>% #arrange the data by the number of movies
      slice(1:10) %>% #select the top 10 directors
      ggplot(aes(x = reorder(director, n), y = n, fill = director)) + #create a bar plot
      geom_bar(stat = "identity") + #use the count of movies as the y variable
      labs(title = "Top 10 Directors with the Most Movies on Netflix", x = "Director", y = "Number of Movies") + #add a title and axis labels
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) #rotate the x axis labels
  })
}
shinyApp(ui = ui, server = server)
```

*Now use ChatGPT, Copilot or previous notes to fix the code*

## That's it! Let's take a break and then move on to working on group projects.
