---
title: "Lab 2 Homework"
author: "Catrinel Berevoescu"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

#### 1. What is a vector in R?  

A vector is a data structure, a way of organizing data in R. It is a collection of items, of a certain data class, combined/concatenated with the function c().  

#### 2. What is a data matrix in R?  

A data matrix is a data structure in R comprised of a collection or series of stacked vectors.  

#### 3. Below are data collected by three scientists (Jill, Steve, Susan in order) measuring temperatures of eight hot springs. Run this code chunk to create the vectors.  

Complete  

```{r}
spring_1 <- c(36.25, 35.40, 35.30)
spring_2 <- c(35.15, 35.35, 33.35)
spring_3 <- c(30.70, 29.65, 29.20)
spring_4 <- c(39.70, 40.05, 38.65)
spring_5 <- c(31.85, 31.40, 29.30)
spring_6 <- c(30.20, 30.65, 29.75)
spring_7 <- c(32.90, 32.50, 32.80)
spring_8 <- c(36.80, 36.45, 33.15)
```

#### 4. Build a data matrix that has the springs as rows and the columns as scientists.  

Creating a new object combining the spring_1 to spring_8 vectors:  

```{r}
temperature_data <- c(spring_1, spring_2, spring_3, spring_4, spring_5, spring_6, spring_7, spring_8)
temperature_data 
```

Creating the data matrix:  

```{r}
temperature_data_matrix <- matrix(temperature_data, nrow = 8, byrow = T)
temperature_data_matrix
```

#### 5. The names of the springs are 1.Bluebell Spring, 2.Opal Spring, 3.Riverside Spring, 4.Too Hot Spring, 5.Mystery Spring, 6.Emerald Spring, 7.Black Spring, 8.Pearl Spring. Name the rows and columns in the data matrix. Start by making two new vectors with the names, then use `colnames()` and `rownames()` to name the columns and rows.  

Creating Row and Column Vectors:  

```{r}
spring_names <- c("Bluebell Spring", "Opal Spring", "Riverside Spring", "Too Hot Spring", "Mystery Spring", "Emerald Spring", "Black Spring", "Pearl Spring") 
spring_names #row names vector
```

```{r}
scientist_names <- c("Jill", "Steve", "Susan")
scientist_names #column names vector
```

Adding Row and Column Names to temperature_data_matrix:  

```{r}
rownames(temperature_data_matrix) <- spring_names #naming rows
```

```{r}
colnames(temperature_data_matrix) <- scientist_names #naming columns
```

```{r}
temperature_data_matrix
```

#### 6. Calculate the mean temperature of all eight springs.

```{r}
mean_temp <- rowMeans(temperature_data_matrix)
mean_temp
```

#### 7. Add this as a new column in the data matrix.  

```{r}
all_temperature_data_matrix <- cbind(temperature_data_matrix, mean_temp)
all_temperature_data_matrix
```

#### 8. Show Susan's value for Opal Spring only.

```{r}
all_temperature_data_matrix[2,3]
```

#### 9. Calculate the mean for Jill's column only.  

```{r}
jill_temp <- all_temperature_data_matrix[,1]
mean(jill_temp)
```

#### 10. Use the data matrix to perform one calculation or operation of your interest.  

Finding and listing the springs with the highest average temperatures (all the hot springs with mean temperatures above the median average hot spring temperature):  

```{r}
all_mean <- all_temperature_data_matrix[,4] #specifying the mean temperatures column
med_all_mean <- median(all_mean) #calculating the median of the mean temperatures column
all_mean[all_mean>=med_all_mean] #finding hot springs with mean temperature above the median
```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  