---
title: "lab 5 warmup"
output: html_document
date: "2024-01-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Warm-up   

```{r}
library(tidyverse)
```

The warm-up will use the fish data from lab 4. These data are from: Gaeta J., G. Sass, S. Carpenter. 2012. Biocomplexity at North Temperate Lakes LTER: Coordinated Field Studies: Large Mouth Bass Growth 2006. Environmental Data Initiative.  [link](https://portal.edirepository.org/nis/mapbrowse?scope=knb-lter-ntl&identifier=267).  

You will need the tidyverse to complete the warm-up- be sure to load the library.  

#### 1. Load the fish data.  

```{r}
fish <- readr::read_csv("data/Gaeta_etal_CLC_data.csv")
```

#### 2. Transform the fish data to only include the variables lakeid and length.  Store this as a new dataframe called "fishlength". 

```{r}
names(fish)
```

```{r}
fishlength <- select(fish, "lakeid", "length")
fishlength
```

#### 3. Filter the `fish` data to include the samples from lake "BO".  

```{r}
glimpse(fish)
```

```{r}
fish_bo <- filter(fish, lakeid == "BO")
fish_bo
```

#### 4. Calculate the mean length of fish from lake "BO".  

```{r}
mean(fish_bo$length)
```

