---
title: "Lab 3 Homework"
author: "Catrinel Berevoescu"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: true
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse   

Complete. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

### Mammals Sleep   

#### 1. For this assignment, we are going to use built-in data on mammal sleep patterns. From which publication are these data taken from? Since the data are built-in you can use the help function in R. The name of the data is `msleep`.   

This data set is taken from the publication by V. M. Savage and G. B. West, A quantitative, theoretical framework for understanding mammalian sleep. Proceedings of the National Academy of Sciences, 104 (3):1051-1056, 2007.

```{r}
?msleep
```

#### 2. Store these data into a new data frame `sleep`.   

```{r}
sleep <- msleep #msleep is already a dataframe
sleep
```

#### 3. What are the dimensions of this data frame (variables and observations)? How do you know? Please show the *code* that you used to determine this below.   

The dimensions of this data frame, as shown by the dim() function below (left value states the number of observations, and the right value states the number of variables), are 83 observations (rows) and 11 variables (columns).  

```{r}
dim(sleep) 
```

#### 4. Are there any NAs in the data? How did you determine this? Please show your code.   

Using the function anyNA(), there are NAs in the data, since running this function came back with "true".  

```{r}
anyNA(sleep)
```

#### 5. Show a list of the column names is this data frame.     

```{r}
names(sleep)
```

#### 6. How many herbivores are represented in the data?     

There are 32 herbivores represented in the data:

```{r}
herbivore_sleep <- filter(sleep, vore == "herbi") #creating new data frame with only herbivores
nrow(herbivore_sleep) #number of rows in herbivore_sleep
```

#### 7. We are interested in two groups; small and large mammals. Let's define small as less than or equal to 19kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.   

Small Mammals Data Frame:  

```{r}
small_mammals <- filter(sleep, bodywt <= 19)
small_mammals
```

Large Mammals Data Frame:  

```{r}
large_mammals <- filter(sleep, bodywt >= 200)
large_mammals
```

#### 8. What is the mean weight for both the small and large mammals?   

Mean Weight for Small Mammals:  

```{r}
mean(small_mammals$bodywt)
```

Mean Weight for Large Mammals:  

```{r}
mean(large_mammals$bodywt)
```

#### 9. Using a similar approach as above, do large or small animals sleep longer on average?   
Mean Sleep Time for Small Mammals:  

```{r}
mean_sleep_small_mammals <- mean(small_mammals$sleep_total) #finding mean of sleep_total column for small_mammal data frame
mean_sleep_small_mammals
```

Mean Sleep Time for Large Mammals:  

```{r}
mean_sleep_large_mammals <- mean(large_mammals$sleep_total) #finding mean of sleep_total column for large_mammal data frame
mean_sleep_large_mammals
```

```{r}
mean_sleep_large_mammals <= mean_sleep_small_mammals #showing that mean sleep time for small mammals is larger than the sleep time for large mammals
```

Since the mean sleep time for small mammals is 11.78644 and the mean sleep time for large mammals is 3.3, this means that small animals sleep longer on average.  

#### 10. Which animal is the sleepiest among the entire dataframe?

To find the sleepiest animal, I am looking for the animal who has the largest sleep_total value:  

```{r}
summary(sleep$sleep_total)
```

The sleepiest animal slept for 19.90 hours (the max value).  

```{r}
sleepiest <- filter(sleep, sleep_total == 19.90)
sleepiest
sleepiest$name
```

Therefore, the sleepiest animal, which has a sleep_total of 19.9, is the little brown bat.  

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   