---
title: "lab_3_warmup"
output: html_document
date: "2024-01-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Warm-up (like a series of midterm questions...hint hint)  

### 1. Build a vector that includes the following height measurements for five plants: plant 1 30.7, plant 2 37.6, plant 3 28.4, plant 4 NA, plant 5 33.2 

```{r}
plant_1_h <-  30.7
plant_2_h <- 37.6
plant_3_h <- 28.4 
plant_4_h <- NA
plant_5_h <- 33.2
plant_heights <- c(plant_1_h, plant_2_h, plant_3_h, plant_4_h, plant_5_h)
```

```{r}
plant_height <- c(30.7, 37.6, 28.4, NA, 33.2)
```

### 2. Build another vector that includes the following mass measurements: plant 1 4, plant 2 5.2, plant 3 3.7, plant 4 NA, plant 5 4.6

```{r}
plant_1_m <-  4
plant_2_m <- 5.2
plant_3_m <- 3.7
plant_4_m <- NA
plant_5_m <- 4.6
plant_masses <- c(plant_1_m, plant_2_m, plant_3_m, plant_4_m, plant_5_m)
```

```{r}
plant_mass <- c(4, 5.2, 3.7, NA, 4.6)
```

### 3. Assemble these vectors into a labeled data matrix with two columns 

Labels for data matrix:  

```{r}
samples <- c("plant_1", "plant_2", "plant_3", "plant_4", "plant_5")
```

```{r}
column_names <- c(plant_heights, plant_masses)
```

```{r}
measured <- c("height", "mass")
```

Combine data for height and weight:  

```{r}
plant_experiment <- c(plant_height, plant_mass)
```

Build the Data Matrix:  

```{r}
plant_data_matrix <- matrix(plant_experiment, nrow = 5, byrow = F) #do reality check for verification here
plant_data_matrix 
```

Naming the columns and rows: 

```{r}
colnames(plant_data_matrix) <- measured
```

```{r}
rownames(plant_data_matrix) <- samples
```

Overall data matrix:  

```{r}
plant_data_matrix
```

### 4. Calculate the mean for height and mass and add them to the data matrix 

Finding the means:  

```{r}
plant_means <- colMeans(plant_data_matrix, na.rm = T) #removing the NAs
plant_means
```

Add this as a new row:  

```{r}
plant_data_matrix_final <- rbind(plant_data_matrix, plant_means)
plant_data_matrix_final
```

