---
title: "HW: dplyr Superhero"
author: "Catrinel Berevoescu"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    toc: true
    toc_float: true
    keep_md: true
---

## Learning Goals  

*At the end of this exercise, you will be able to:*    
1. Develop your dplyr superpowers so you can easily and confidently manipulate dataframes.  
2. Learn helpful new functions that are part of the `janitor` package.  

## Instructions   

For the second part of lab today, we are going to spend time practicing the dplyr functions we have learned and add a few new ones. This lab doubles as your homework. Please complete the lab and push your final code to GitHub.  

## Load the libraries    

```{r}
library("tidyverse")
library("janitor")
```

## Load the superhero data   

These are data taken from comic books and assembled by fans. The include a good mix of categorical and continuous data.  Data taken from: https://www.kaggle.com/claudiodavi/superhero-set  

Check out the way I am loading these data. If I know there are NAs, I can take care of them at the beginning. But, we should do this very cautiously. At times it is better to keep the original columns and data intact.   

```{r}
superhero_info <- read_csv("data/heroes_information.csv", na = c("", "-99", "-"))
superhero_powers <- read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))
```

## Data tidy.   

#### 1. Some of the names used in the `superhero_info` data are problematic so you should rename them here. Before you do anything, first have a look at the names of the variables. You can use `rename()` or `clean_names()`.    

```{r}
names(superhero_info)
```

```{r}
superhero_info <- clean_names(superhero_info)
superhero_powers <- clean_names(superhero_powers)
```

## `tabyl`
The `janitor` package has many awesome functions that we will explore. Here is its version of `table` which not only produces counts but also percentages. Very handy! Let's use it to explore the proportion of good guys and bad guys in the `superhero_info` data.  

```{r}
tabyl(superhero_info, alignment)
```

#### 1. Who are the publishers of the superheros? Show the proportion of superheros from each publisher. Which publisher has the highest number of superheros?  

Publishers of the Superheroes:   

```{r}
table(superhero_info$publisher)
```

Proportions of Superheroes from each Publisher:   

```{r}
superhero_info %>% 
        tabyl(publisher) %>% 
        arrange(desc(n))
```

##### The publishers of the superheroes are shown above (24 total). The publisher with the highest number of superheroes appears to be Marvel comics with 388 superheroes.    

#### 2. Notice that we have some neutral superheros! Who are they? List their names below.  

```{r}
neutral_superheroes <- superhero_info %>% 
  select(name, alignment) %>% 
  filter(alignment == "neutral") #filtering for neutral superheroes
neutral_superheroes
```

List of Neutral Superheroes:   

```{r}
tabyl(neutral_superheroes, name) 
```

##### There are 24 neutral superheroes represented in the superhero_info data frame.    

## `superhero_info`   

#### 3. Let's say we are only interested in the variables name, alignment, and "race". How would you isolate these variables from `superhero_info`?    

I selected the three desired variables name, alignment, and "race" with the select() function.

```{r}
superhero_info %>% 
  select(name, alignment, race) #isolating desired variables with the select() function
```

## Not Human    

#### 4. List all of the superheros that are not human.    

```{r}
nonhuman_superheroes <- superhero_info %>% 
        filter(race != "Human") #defining superheroes as human only if their "race" is described exclusively as "human"
nonhuman_superheroes
```

Listing the Superheroes that are Not Human:    

```{r}
tabyl(nonhuman_superheroes, name)
```

## Good and Evil   

#### 5. Let's make two different data frames, one focused on the "good guys" and another focused on the "bad guys".    

Good Guys Data Frame:   

```{r}
good_guys <- filter(superhero_info, alignment == "good")
good_guys
```

Bad Guys Data Frame:   

```{r}
bad_guys <- filter(superhero_info, alignment == "bad")
bad_guys
```

#### 6. For the good guys, use the `tabyl` function to summarize their "race".    

```{r}
tabyl(good_guys, race) #summarizing the good guys data frame by "race"
```


#### 7. Among the good guys, Who are the Vampires?    

```{r}
filter(good_guys, race == "Vampire")
```

##### The Vampires among the good guys are Angel and Blade.    

#### 8. Among the bad guys, who are the male humans over 200 inches in height?    

```{r}
bad_guys %>% 
  filter(race == "Human", gender == "Male", height > 200) #selecting male humans over 200 inches in height
```

##### Among the bad guys, the male humans over 200 inches in height are Bane, Doctor Doom, Kingpin, Lizard, and Scorpion.    

#### 9. Are there more good guys or bad guys with green hair?    

Good Guys with Green Hair:   

```{r}
good_guys %>% 
  filter(hair_color == "Green")
```

```{r}
good_guys %>% 
        filter(hair_color == "Green") %>% 
        dim()
```

Bad Guys with Green Hair:    

```{r}
bad_guys %>% 
  filter(hair_color == "Green")
```

```{r}
bad_guys %>% 
        filter(hair_color == "Green") %>% 
        dim()
```

##### Since there are 7 good guys with green hair, and only 1 bad guy with green hair, there are more good guys with green hair.    

#### 10. Let's explore who the really small superheros are. In the `superhero_info` data, which have a weight less than 50? Be sure to sort your results by weight lowest to highest.  

```{r}
superhero_info %>% #searching within the superhero_info data frame
  filter(weight < 50) %>% #filtering for superheroes that weigh less than 50
  arrange(weight) #arranging the results from lowest to highest weight
```

For ease of viewing, I have selected only the `name` and `weight` variables below:    

```{r}
superhero_info %>% #searching within the superhero_info data frame
        filter(weight < 50) %>% #filtering for superheroes that weigh less than 50 
        select(name, weight) %>% #selecting only the variables name and weight
        arrange(weight) #arranging the results from lowest to highest weight
```

#### 11. Let's make a new variable that is the ratio of height to weight. Call this variable `height_weight_ratio`.     

```{r}
superhero_info %>% 
        mutate(height_weight_ratio = height / weight) #adding a new column with the height to weight ratio
```

For a more clear visual of the `height_weight_ratio` variable:    

```{r}
superhero_info %>% 
        mutate(height_weight_ratio = height / weight) %>% 
        select(name, height_weight_ratio)
```

#### 12. Who has the highest height to weight ratio?  

```{r}
superhero_info %>% 
        mutate(height_weight_ratio = height / weight) %>% 
        arrange(desc(height_weight_ratio))
```

For ease of viewing, I have selected only the `name` and `height_weight_ratio` variables below:   

```{r}
superhero_info %>% 
        mutate(height_weight_ratio = height / weight) %>% 
        select(name, height_weight_ratio) %>% 
        arrange(desc(height_weight_ratio))
```

##### Thus, the superhero with the highest height to weight ratio is Groot (175.25000000).   

## `superhero_powers`    

Have a quick look at the `superhero_powers` data frame.  

```{r}
superhero_powers <- clean_names(superhero_powers) #tidying up the superhero_powers data frame
```

```{r}
glimpse(superhero_powers) #looking at the superhero_powers data frame
```

#### 13. How many superheros have a combination of agility, stealth, super_strength, stamina?    

```{r}
specific_superheroes <- superhero_powers %>% 
        filter(agility == "TRUE", stealth == "TRUE", super_strength == "TRUE", stamina == "TRUE") %>%  #filtering for superheroes that have this combination of powers (as in, all of powers are "TRUE" for each superhero)
        select(hero_names, agility, stealth, super_strength, stamina) #selecting only the specific combination of powers for ease of viewing
specific_superheroes
```

```{r}
dim(specific_superheroes)
```

##### Thus, there are 40 superheroes that have the combination of agility, stealth, super_strength, stamina powers.    

## Your Favorite   

#### 14. Pick your favorite superhero and let's see their powers!  

My favorite superhero is Catwoman:

```{r}
catwoman_powers <- superhero_powers %>% 
        filter(hero_names == "Catwoman") %>% 
        select_if(.==TRUE) 
catwoman_powers
```

```{r}
names(catwoman_powers)
```

##### Thus, Catwoman's powers include agility, stealth, marksmanship, animal attributes, stamina, dexterity, peak human condition, and empathy.    

#### 15. Can you find your hero in the superhero_info data? Show their info!  

I could find my hero, Catwoman, in the superhero_info data:    

```{r}
superhero_info %>% 
        filter(name == "Catwoman") #filtering by the name Catwoman to show only her info
```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  