---
title: "dplyr Superhero"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    toc: yes
    toc_float: yes
    keep_md: yes
---

## Learning Goals  

*At the end of this exercise, you will be able to:*    
1. Develop your dplyr superpowers so you can easily and confidently manipulate dataframes.  
2. Learn helpful new functions that are part of the `janitor` package.  

## Instructions    

For the second part of lab today, we are going to spend time practicing the dplyr functions we have learned and add a few new ones. This lab doubles as your homework. Please complete the lab and push your final code to GitHub.  

## Load the libraries   

```{r}
library("tidyverse")
library("janitor")
```

## Load the superhero data   

These are data taken from comic books and assembled by fans. The include a good mix of categorical and continuous data.  Data taken from: https://www.kaggle.com/claudiodavi/superhero-set  

Check out the way I am loading these data. If I know there are NAs, I can take care of them at the beginning. But, we should do this very cautiously. At times it is better to keep the original columns and data intact.  

```{r}
superhero_info <- read_csv("data/heroes_information.csv", na = c("", "-99", "-")) #told what to interpret as NAs in the read_csv(), but don't add this to every data file
superhero_powers <- read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))
```

## Data tidy    

#### 1. Some of the names used in the `superhero_info` data are problematic so you should rename them here. Before you do anything, first have a look at the names of the variables. You can use `rename()` or `clean_names()`.    

```{r}
names(superhero_info)
```

```{r}
superhero_info <- clean_names(superhero_info)
```

## `tabyl`   

The `janitor` package has many awesome functions that we will explore. Here is its version of `table` which not only produces counts but also percentages. Very handy! Let's use it to explore the proportion of good guys and bad guys in the `superhero_info` data.  

```{r}
tabyl(superhero_info, alignment)
```

#### 1. Who are the publishers of the superheros? Show the proportion of superheros from each publisher. Which publisher has the highest number of superheros?  

```{r}
table(superhero_info$publisher)
```
##### The publishers of the superheroes are shown above (24 total). The publisher with the highest number of superheroes appears to be Marvel comics with 388 superheroes.

#### 2. Notice that we have some neutral superheros! Who are they? List their names below.  

```{r}
superhero_info %>% 
  select(name, alignment) %>% 
  filter(alignment == "neutral")
```

##### There are 24 neutral superheroes represented in the superhero_info data frame.

## `superhero_info`  

#### 3. Let's say we are only interested in the variables name, alignment, and "race". How would you isolate these variables from `superhero_info`?

I selected the three desired variables name, alignment, and "race" with the select() function.

```{r}
superhero_info %>% 
  select(name, alignment, race)
```


## Not Human   

#### 4. List all of the superheros that are not human.

```{r}
superhero_info %>% 
  filter(race != "Human", race != "Human / Radiation", race != "Human / Clone", race != "Human / Cosmic", race != "Human / Altered")

```

## Good and Evil    

#### 5. Let's make two different data frames, one focused on the "good guys" and another focused on the "bad guys".   

Good Guys Data Frame:

```{r}
good_guys <- filter(superhero_info, alignment == "good")
good_guys
```

Bad Guys Data frame:

```{r}
bad_guys <- filter(superhero_info, alignment == "bad")
bad_guys
```


#### 6. For the good guys, use the `tabyl` function to summarize their "race".

```{r}
tabyl(good_guys, race)
```


#### 7. Among the good guys, Who are the Vampires?

```{r}
filter(good_guys, race == "Vampires")
```

##### There are not Vampires among the good guys.   

#### 8. Among the bad guys, who are the male humans over 200 inches in height?   

```{r}
bad_guys %>% 
  filter(race == "Human", gender == "Male", height > 200) #selecting male humans over 200 inches in height
```
##### Among the bad guys, the male humans over 200 inches in height are Bane, Doctor Doom, Kingpin, Lizard, and Scorpion.   

#### 9. Are there more good guys or bad guys with green hair?  

Good Guys with Green Hair:

```{r}
good_guys %>% 
  filter(hair_color == "Green")
```

Bad Guys with Green Hair:

```{r}
bad_guys %>% 
  filter(hair_color == "Green")
```

##### Since there are 7 good guys with green hair, and only 1 bad guy with green hair, there are more good guys with green hair.

#### 10. Let's explore who the really small superheros are. In the `superhero_info` data, which have a weight less than 50? Be sure to sort your results by weight lowest to highest.  

```{r}
superhero_info %>% 
  filter(weight < 50) %>% 
  arrange(weight)
```

#### 11. Let's make a new variable that is the ratio of height to weight. Call this variable `height_weight_ratio`.    

#### 12. Who has the highest height to weight ratio?  

## `superhero_powers`    

Have a quick look at the `superhero_powers` data frame.  

#### 13. How many superheros have a combination of agility, stealth, super_strength, stamina?

## Your Favorite    

#### 14. Pick your favorite superhero and let's see their powers!  

#### 15. Can you find your hero in the superhero_info data? Show their info!  

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  