---
title: "lab6_warmup"
output: html_document
date: "2024-01-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Warm-up  

### 1. Load the `bison.csv` data.   (library calls)

```{r}
library(tidyverse) #make sure to embed the library calls in every bit of data so other people can easily run your code
library(janitor)
```
```{r}
bison <- read_csv("data/bison.csv")
bison <- clean_names(bison)
```

### 2. What are the dimesions and structure of the data?

The bison data frame is a data frame (each variable can be a different data class, see data classes in the <> below).

```{r}
glimpse(bison) #taking a look at the data (dbl = double floating point integers)
```

Dimensions of Bison:

There are 8 variables (columns) and 8325 observations (rows).

```{r}
dim(bison)
```


### 3. We are only interested in code, sex, weight, year of birth. Restrict the data to these variables and store the dataframe as a new object.  

```{r}
names(bison)
```

New, restricted Bison dataframe:

```{r}
restricted_bison <- select(bison, "data_code", "animal_code", "animal_sex", "animal_weight", "animal_yob")
restricted_bison
```

Another method, using pipes (from now on, use pipes):

```{r}
bison_new <- bison %>% 
  select(animal_code, animal_sex, animal_weight, animal_yob)
```


### 4. Pull out the animals born between 1980-1990.   

```{r}
specific_animal_birthdays <- filter(bison, between(animal_yob, 1980, 1990)) #using between() may be more elegant, but can use whatever you want
specific_animal_birthdays 
```

Another method:

```{r}
bison_new %>%
  filter(animal_yob >= 1980 & animal_yob <= 1990) #use an & here
```


### 5. How many male and female bison are represented between 1980-1990?

Filtering specific_animal_birthdays by females:

```{r}
females <- filter(specific_animal_birthdays, animal_sex=="F")
females
```

There are 414 female bison in the specific_animal_birthdays data frame:

```{r}
dim(females)
```

Filtering specific_animal_birthdays by males:

```{r}
males <- filter(specific_animal_birthdays, animal_sex=="M")
males
```

There are 21 male bison in the specific_animal_birthdays dataframe: 

```{r}
dim(males)
```

Another method:

```{r}
male_bison <- bison_new %>% #first pull out males
  filter(animal_yob >= 1980 & animal_yob <= 1990) %>% 
  filter(animal_sex=="M") #could change this to filter out for females
male_bison

table(male_bison$animal_sex) #another way to find the counts (observations) of male bison
```

```{r}
female_bison <- bison_new %>% #first pull out females
  filter(animal_yob >= 1980 & animal_yob <= 1990) %>% 
  filter(animal_sex=="F") 
female_bison

table(female_bison$animal_sex) #another way to find the counts (observations) of female bison
```

### 6. Between 1980-1990, were males or females larger on average?    

Female Mean:    

```{r}
average_female_weight <- mean(females$animal_weight)
average_female_weight
```

Male Mean:   

```{r}
average_male_weight <- mean(males$animal_weight)
average_male_weight
```

```{r}
average_male_weight > average_female_weight
```

Thus, between 1980-1990, males were larger on average.