---
title: "Lab 4 Homework"
author: "Catrinel Berevoescu"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions   

Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse  

Complete  

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Data  

For the homework, we will use data about vertebrate home range sizes. The data are in the class folder, but the reference is below.  

**Database of vertebrate home range sizes.**  

Reference: Tamburello N, Cote IM, Dulvy NK (2015) Energy and the scaling of animal space use. The American Naturalist 186(2):196-211. http://dx.doi.org/10.1086/682070.  
Data: http://datadryad.org/resource/doi:10.5061/dryad.q5j65/1  

#### 1. Load the data into a new object called `homerange`.   

```{r}
homerange <- read.csv("data/Tamburelloetal_HomeRangeDatabase.csv")
```

```{r}
#install.packages("janitor") #I don't want to reinstall this package so I added a "#"
```

```{r}
library(janitor)
```

```{r}
homerange <- clean_names(homerange) #tidying up data frame
```

#### 2. Explore the data. Show the dimensions, column names, classes for each variable, and a statistical summary. Keep these as separate code chunks.  

Dimensions:

569 rows (observations), 24 columns (variables)

```{r}
dim(homerange)
```

Column Names:

```{r}
names(homerange)
```

Classes for Each Variable:  

```{r}
glimpse(homerange) #see classes in "<>"
```

```{r}
sapply(homerange, class) #a simplified summary of the data classes for each variable; the full data class name is provided in quotes for each variable
```

Statistical Summary (provided for numerical variables):   

```{r}
summary(homerange)
```

#### 3. Change the class of the variables `taxon` and `order` to factors and display their levels.  

Taxon:

```{r}
taxon <- as.factor(homerange$taxon) #changing the class of the variables 'taxon' to factors
levels(taxon) #displaying levels of 'taxon'
```

Order:

```{r}
order <- as.factor(homerange$order) #changing the class of the variables 'order' to factors
levels(order) #displaying levels of 'order'
```

#### 4. What taxa are represented in the `homerange` data frame? Make a new data frame `taxa` that is restricted to taxon, common name, class, order, family, genus, species.  

The following taxa, where taxa are generic names for taxonomic groupings such as species, genus, and family, are represented in the homerange data frame.

Data frame 'taxa':

```{r}
taxa <- select(homerange, "taxon", "common_name", "class", "order", "family", "genus", "species")
taxa
```

##### There are 9 taxon represented in the data frame 'homerange': birds, lake fishes, lizards, mammals, marine fishes, river fishes, snakes, tortoises, and turtles.

```{r}
levels(taxon) #the taxon represented in the data frame homerange
```


#### 5. The variable `taxon` identifies the common name groups of the species represented in `homerange`. Make a table the shows the counts for each of these `taxon`.  

Counts for each of the 'taxon':

```{r}
table(homerange$taxon) #counts
```

Table for 'taxon' counts:

```{r}
taxon_table <- data.frame(table(taxon))
taxon_table <- rename(taxon_table, counts=Freq)
taxon_table #table
```

#### 6. The species in `homerange` are also classified into trophic guilds. How many species are represented in each trophic guild.  

##### There are two trophic guilds in the 'homerange' data frame, there are 342 species represented in the carnivore category, and 227 species represented in the herbivore category.

```{r}
table(homerange$trophic_guild)
```

#### 7. Make two new data frames, one which is restricted to carnivores and another that is restricted to herbivores.   

Note: completed in lab day 6 and as homework.  

Carnivores Data Frame:   

```{r}
carnivores <- filter(homerange, trophic_guild == "carnivore")
```

Herbivores Data Frame:   

```{r}
herbivores <- filter(homerange, trophic_guild == "herbivore")
```

#### 8. Do herbivores or carnivores have, on average, a larger `mean.hra.m2`? Remove any NAs from the data.   

Note: completed in lab day 6 and as homework.   

```{r}
carnivores_mean_hra_m2 <- mean(carnivores$mean_hra_m2, rm.NA = T)
carnivores_mean_hra_m2
```

```{r}
herbivores_mean_hra_m2 <- mean(herbivores$mean_hra_m2, rm.NA = T)
herbivores_mean_hra_m2
```

```{r}
herbivores_mean_hra_m2 > carnivores_mean_hra_m2
```

##### Herbivores have, on average, a larger `mean_hra_m2` than carnivores.

#### 9. Make a new dataframe `owls` that is limited to the mean mass, log10 mass, family, genus, and species of owls in the database. Which is the smallest owl? What is its common name? Do a little bit of searching online to see what you can learn about this species and provide a link below   

Note: completed in lab day 6 and as homework.   

Owls Data Frame:

```{r}
owls_selected <- filter(homerange, order == "strigiformes")
owls <- select(owls_selected, "family", "genus", "species", "mean_mass_g", "log10_mass")
owls
```

```{r}
summary(owls$mean_mass_g)
```

The smallest owl weighs 61.32 grams, the minimum mean_mass_g value.

Identifying the Smallest Owl:   

```{r}
filter(homerange, mean_mass_g == 61.32)
```

##### The smallest owl is Glaucidium passerinum. Its common name is the Eurasian pygmy owl.

##### The Eurasian pygmy owl is the smallest owl of Europe. It is native to the central area of the Palearctic region, and lives in coniferous forests. Due to their large feet and legs, they can catch prey, which are typically small mammals and birds, up to their own size. Their typical lifespan is 6 years.    

##### More about the [Eurasian pygmy owl](https://animaldiversity.org/accounts/Glaucidium_passerinum/)

#### 10. As measured by the data, which bird species has the largest homerange? Show all of your work, please. Look this species up online and tell me about it!.    

Note: completed in lab day 6 and as homework.   

```{r}
homerange %>% #searching within the homerange data frame
  filter(taxon == "birds") %>% #filtering for bird species
  select(species, mean_hra_m2) %>% #selecting the desired columns
  arrange(desc(mean_hra_m2)) #arranging the data to have the largest homerange value shown first

```

The bird species with the largest homerange as found above is the cheriway, with a homerange of 241000000.00 square meters.  

```{r}
filter(homerange, species == "cheriway") #filtering for the species with the largest homerange
```


##### Thus, the bird species with the largest homerange is the Caracara cheriway, common name caracara.

##### The Caracara cheriway is found along the Mexican-American Border. They are opportunistic feeders that eat small animals and live in open countrysides. They are also quite aggressive birds, and will chase other birds and scavengers, like vultures, away from prey and road kills. The name "caracara" comes from from the sound of their call. Appearance-wise they have colorful faces with black crests and bellies.

##### More about the [Caracara cheriway](https://animaldiversity.org/accounts/Caracara_cheriway/).   [Caracara cheriway - The Crested Caracara](https://www.audubon.org/field-guide/bird/crested-caracara).  

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   