---
title: "Lab 7 Homework"
author: "Catrinel Berevoescu"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions   

Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries     

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
```

For this assignment we are going to work with a large data set from the [United Nations Food and Agriculture Organization](http://www.fao.org/about/en/) on world fisheries. These data are pretty wild, so we need to do some cleaning. First, load the data.  

Load the data `FAO_1950to2012_111914.csv` as a new object titled `fisheries`.   

```{r}
fisheries <- readr::read_csv(file = "data/FAO_1950to2012_111914.csv")
```

#### 1. Do an exploratory analysis of the data (your choice). What are the names of the variables, what are the dimensions, are there any NA's, what are the classes of the variables?     

Names of the Variables:   

```{r}
names(fisheries)
```

Dimensions:    

There are 71 columns (variables) and 17692 rows (observations) in the fisheries data.   

```{r}
dim(fisheries)
```

Since the anyNA() function returned "true", there ARE NAs in the file:   

```{r}
anyNA(fisheries)
```

Data Classes of the Variables:   

```{r}
glimpse(fisheries)
```

#### 2. Use `janitor` to rename the columns and make them easier to use. As part of this cleaning step, change `country`, `isscaap_group_number`, `asfis_species_number`, and `fao_major_fishing_area` to data class factor.    

```{r}
fisheries <- clean_names(fisheries)
```

```{r}
names(fisheries)
```

```{r}
isscaap_group_number <- as.factor(fisheries$isscaap_group_number)
asfis_species_number <- as.factor(fisheries$asfis_species_number)
fao_major_fishing_area <- as.factor(fisheries$fao_major_fishing_area)
```

We need to deal with the years because they are being treated as characters and start with an X. We also have the problem that the column names that are years actually represent data. We haven't discussed tidy data yet, so here is some help. You should run this ugly chunk to transform the data for the rest of the homework. It will only work if you have used janitor to rename the variables in question 2!    

```{r}
fisheries_tidy <- fisheries %>% 
  pivot_longer(-c(country,common_name,isscaap_group_number,isscaap_taxonomic_group,asfis_species_number,asfis_species_name,fao_major_fishing_area,measure),
               names_to = "year",
               values_to = "catch",
               values_drop_na = TRUE) %>% 
  mutate(year= as.numeric(str_replace(year, 'x', ''))) %>% 
  mutate(catch= str_replace(catch, c(' F'), replacement = '')) %>% 
  mutate(catch= str_replace(catch, c('...'), replacement = '')) %>% 
  mutate(catch= str_replace(catch, c('-'), replacement = '')) %>% 
  mutate(catch= str_replace(catch, c('0 0'), replacement = ''))

fisheries_tidy$catch <- as.numeric(fisheries_tidy$catch)
```

#### 3. How many countries are represented in the data? Provide a count and list their names.   

##### There are 204 unique Countries represented in the data:    

```{r}
fisheries_tidy %>% 
        summarize(n_countries = n_distinct(country))
```

List of Country Names:   

```{r}
fisheries_tidy %>% 
        count(country)
```

#### 4. Refocus the data only to include country, isscaap_taxonomic_group, asfis_species_name, asfis_species_number, year, catch.   

```{r}
fisheries_refocused <- fisheries_tidy %>% 
        select(country, isscaap_taxonomic_group, asfis_species_name, asfis_species_number, year, catch)
fisheries_refocused
```

#### 5. Based on the asfis_species_number, how many distinct fish species were caught as part of these data?   

##### There were 1551 distinct fish species caught as part of the data:  

```{r}
fisheries_refocused %>% 
        summarize(n_fish_species = n_distinct(asfis_species_number))
```

#### 6. Which country had the largest overall catch in the year 2000?    

```{r}
fisheries_refocused %>% 
        filter(year == 2000) %>% #filtering by the year 2000
        group_by(country) %>% #grouping the data by the categorical variable country
        summarize(overall_catch = sum(catch, na.rm = TRUE)) %>% #calculating the overall catch values
        arrange(desc(overall_catch)) #arranging the data to show the countries that had the largest ovcerall catches in the year 2000 first
```

##### Thus, the country with the largest overall catch in the year 2000 was China, with 25899 fish caught.    

#### 7. Which country caught the most sardines (_Sardina pilchardus_) between the years 1990-2000?    

```{r}
fisheries_refocused %>% 
        filter(year >= 1990, year <= 2000, asfis_species_name == "Sardina pilchardus") %>% #filtering by the time frame and sardines
        group_by(country) %>% #grouping the data by the categorical variable country
        summarize(total_sardines = sum(catch, na.rm = T)) %>% #finding total number of sardines caught for each country within the time frame
        arrange(desc(total_sardines)) #arranging the data to show the countries that caught the most sardines first
```

##### Thus, the country that caught the most sardines (_Sardina pilchardus_) between the years 1990-2000 was Morocco, with 7470 sardines caught.    

#### 8. Which five countries caught the most cephalopods between 2008-2012?    

```{r}
fisheries_refocused %>% 
        filter(year >= "2008", year <= "2012", asfis_species_name == "Cephalopoda") %>% #filtering for the time frame and cephalopods
        group_by(country) %>% #grouping the data by the categorical variable country
        summarize(total_cephalopods = sum(catch, na.rm = T)) %>% #adding together cephalopods caught between 2008-2012 to find total cephalopods caught by each country during this time period
        arrange(desc(total_cephalopods)) %>% #arranging such that the countries that caught the most cephalopods is shown at the top
        head()
```

##### The five countries that caught the most cephalopods between 2008-2012 are India (570), China (257), Spain (198), Algeria (162), and France (101).    

#### 9. Which species had the highest catch total between 2008-2012? (hint: Osteichthyes is not a species).   

```{r}
fisheries_refocused %>% 
        filter(year >= "2008", year <= "2012") %>% #filtering by the time frame
        group_by(asfis_species_name) %>% #grouping the data by the categorical variable asfis_species_name
        summarize(total_catch = sum(catch, na.rm = T)) %>% #adding together fish of a certain species caught between 2008-2012 to find total fish caught during this time period
        arrange(desc(total_catch)) #arranging such that the species that was caught the most is shown first
```

```{r}
fisheries_refocused %>% 
        filter(year >= "2008", year <= "2012") %>% #filtering by the time frame
        group_by(asfis_species_number) %>% #grouping the data by the categorical variable asfis_species_number
        summarize(total_catch = sum(catch, na.rm = T)) %>% #adding together fish of a certain species caught between 2008-2012 to find total fish caught during this time period
        arrange(desc(total_catch)) #arranging such that the species that was caught the most is shown first
```

##### Thus the species that is caught the most is Pisces miscellanea (species number = 199XXXXXXX010) from the Osteichthyes, which is a superclass of vertebrate animals that are more commonly known as "bony fish".    

#### 10. Use the data to do at least one analysis of your choice. 

##### I want to find the species that was caught the most on average over the entire time period represented in the data set:   

```{r}
fisheries_refocused %>% 
        group_by(asfis_species_number) %>% 
        summarize(average_catch = mean(catch, na.rm = T)) %>% 
        arrange(desc(average_catch))
```

```{r}
fisheries_refocused %>% 
        group_by(asfis_species_name) %>% 
        summarize(average_catch = mean(catch, na.rm = T)) %>% 
        arrange(desc(average_catch))
```

##### The fish species that was caught the most on average in the represented time period is species number 1210600208, Engraulis ringens.   

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   